---
title: "R Notebook"
output: html_notebook
---

```{r Libraries, message=FALSE, warning=FALSE}
library(osmdata)
library(sf)
library(dplyr)
library(ggplot2)
library(lwgeom)
library(tidyr)
library(stringr)
```

```{r}
### define example bbox
lon_min <- 6.5 # xmin
lon_max <- 13 # xmax
lat_min <- 53.5 # ymin
lat_max <- 58.5 # ymax

bb <- matrix(c(lon_min,lon_max,lat_min,lat_max), byrow = TRUE, nrow = 2, ncol = 2, dimnames = list(c('x','y'),c('min','max'))) 
bb

```

```{r}
### get "water" that is not sea as polygons
# water <- opq(bb) %>%
#   add_osm_feature(key = "natural", value = "water") %>%
#   osmdata_sf()
```


```{r}
### get sea & land as polygons
# 1. get coastline (as line)
coast <- opq(bb) %>%
  add_osm_feature(key = "natural", value = "coastline") %>%
  osmdata_sf()

```

```{r message=FALSE, warning=FALSE}
polygon_Map <- ggplot() +
  geom_sf(
    data = coast$osm_polygons,
    fill = "azure",
    color = "steelblue"
  ) +
  coord_sf(xlim = c(lon_min, lon_max), 
           ylim = c(lat_min, lat_max),
           expand = FALSE) +
    theme(axis.line = element_line(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=.3),
        panel.background = element_rect(fill = "white"))

print(polygon_Map)

scale_factor = 10

## Save as PNG file
ggsave("Gifs/Map1_Polygons_%03d.png",
       plot = last_plot(), 
       device = "png",
       scale = scale_factor,
       units="mm", 
       limitsize = FALSE,
       dpi=300)
```

```{r message=FALSE, warning=FALSE}
line_Map <- ggplot() +
  geom_sf(
    data = coast$osm_lines,
    fill = "azure",
    color = "steelblue",
    linewidth=.3
  ) + 
  coord_sf(xlim = c(lon_min, lon_max), 
           ylim = c(lat_min, lat_max),
           expand = FALSE) +
    theme(axis.line = element_line(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=.3),
        panel.background = element_rect(fill = "white"))

print(line_Map)

scale_factor = 10

## Save as PNG file
ggsave("Gifs/Map1_Line_%03d.png",
       plot = last_plot(), 
       device = "png",
       scale = scale_factor,
       units="mm", 
       limitsize = FALSE,
       dpi=300)

```


```{r message=FALSE, warning=FALSE}
line_and_polygon_Map <- ggplot() +
    geom_sf(
    data = coast$osm_polygons,
    fill = "red",
    color = "green"#,
  #  linewidth=.3
    ) +
  geom_sf(
    data = coast$osm_lines,
    fill = "azure",
    color = "steelblue",
    linewidth=.3
  ) + 
  coord_sf(xlim = c(lon_min, lon_max), 
           ylim = c(lat_min, lat_max),
           expand = FALSE) +
    theme(axis.line = element_line(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=.3),
        panel.background = element_rect(fill = "white"))

print(line_and_polygon_Map)

scale_factor = 10

## Save as PNG file
ggsave("Gifs/Map1_Line_and_Polygon_%03d.png",
       plot = last_plot(), 
       device = "png",
       scale = scale_factor,
       units="mm", 
       limitsize = FALSE,
       dpi=300)

```

There are a small number of islets that are not included in the osm_lines. So merge all of the polygons (*osm_polygons*) into lines (*osm_lines*) with the function **osm_poly2line**.

```{r}
# Merge all of the osm_polygon objects in coast into osm_lines. The osm_polygons are retained but the are also repeated as osm_lines
coast_all <- osm_poly2line(coast)

dim(coast$osm_lines)
dim(coast_all$osm_lines)
```

However to create a map with sea and land in colour we need to convert osm_lines into polygons. 
This is done by creating a rectangle/polygon which is the size of the bounding box (bbox) and then spliting it via the coastline (**coast_all$osm_lines**). 
```{r}
# Get overall rectangle for bbox
bb_rect <- data.frame(
  lat = c(lat_min, lat_max),
  lon = c(lon_min, lon_max)
) %>% 
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>% 
  st_bbox() %>% 
  st_as_sfc()

```


```{r}
# Split overall rectangle for bbox via coastline
bb_rect_split <- bb_rect %>% 
  st_split(coast_all$osm_lines) %>% 
  st_collection_extract("POLYGON")
```

```{r}
# Extract the split parts
sea <- bb_rect_split[1]
```


```{r message=FALSE, warning=FALSE}
Map1 <- ggplot() +
    geom_sf(
    data = sea,
    fill = "aliceblue",
    color = "skyblue1", # outline
    linewidth=.3
    ) +
  coord_sf(xlim = c(lon_min, lon_max), 
           ylim = c(lat_min, lat_max),
           expand = FALSE) +
    theme(axis.line = element_line(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=.3),
        panel.background = element_rect(fill = "white"))

print(Map1)

scale_factor = 10

## Save as PNG file
ggsave("Gifs/Map1_%03d.png",
       plot = last_plot(), 
       device = "png",
       scale = scale_factor,
       units="mm", 
       limitsize = FALSE,
       dpi=300)

```

```{r}
# Read in file containing towns, seas and countries
Map_coordinates <- readxl::read_excel("Data/Map_coordinates.xlsx", 
                           sheet = "Sheet1", range = "A1:D18",
                           col_types = c("text", "numeric", "numeric", "text")) %>%
    # Add a new column "Mark" which is set to Label if the Type is Sea or Country and otherwise set to Point
  dplyr::mutate(Mark = case_when(Type %in% c("Sea", "Country", "Area") ~ "Label",
                                 TRUE ~ "Point")) %>%
  # Add a new column "Fontsize" which is set to 4 if Type is Country, 
  # 3 if Sea and 2.5 if any other value
  dplyr::mutate(FontSize = case_when(
    Type == "Country" ~ 4,
    Type == "Sea" ~ 3,
    TRUE ~ 3))  %>%
  dplyr::mutate(FontType = case_when(
    Type == "Country" ~ "bold",
    Type == "Sea" ~ "bold.italic",
    Type == "Area" ~ "italic",
    TRUE ~ "plain"))  %>%
  dplyr::mutate(FontColour = case_when(
    Type == "Country" ~ "black",
    Label %in% c("Puttgarden","Rødbyhavn") ~ "black",
    Type == "Sea" ~ "steelblue",
    TRUE ~ "darkgray"))  %>%
    dplyr::mutate(PointSize = case_when(
    Label %in% c("Puttgarden","Rødbyhavn") ~ 1.5,
    TRUE ~ 1))  %>%
    # Add a new column "NudgeX" which is set to 0 
  dplyr::mutate(NudgeX = case_when(
    Label %in% c("Puttgarden","Rødbyhavn") ~ -.36,
    Type == "Country" ~ 0,
    Type == "Sea" ~ 0,
    TRUE ~ 0)) %>%
    # Add a new column "NudgeX" which is set to 0 
  dplyr::mutate(NudgeY = case_when(
    Label %in% c("Puttgarden","Rødbyhavn") ~ 0,
    Type == "Country" ~ 0,
    Type == "Sea" ~ 0,
    TRUE ~ -.05))

  
head(Map_coordinates,3)
```

```{r}
Labels <- Map_coordinates %>%
  sf::st_as_sf(coords = c("Long", "Lat"), crs = 4326, agr = "constant")

Points <- dplyr::filter(Map_coordinates, !Mark == "Label") %>%
  sf::st_as_sf(coords = c("Long", "Lat"), crs = 4326, agr = "constant")

```



```{r message=FALSE, warning=FALSE}
Map2 <- ggplot() +
    geom_sf(
    data = sea,
    fill = "aliceblue",
    color = "skyblue1", # outline
    linewidth=.2
    ) +
    geom_sf(data = Points, size = Points$PointSize, shape = 21, fill = "darkred") +
    geom_sf_text(data = Labels, aes(label = Label, lineheight = .85), 
               size = Labels$FontSize,
               colour = Labels$FontColour,
               fontface = Labels$FontType,
               nudge_x = Labels$NudgeX,
               nudge_y = Labels$NudgeY) +
  coord_sf(xlim = c(lon_min, lon_max), 
           ylim = c(lat_min, lat_max-2.5),
           expand = FALSE) +
    theme(axis.line = element_line(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.title =element_blank(),  # hides the X and y axis labels
        panel.border = element_rect(colour = "black", fill=NA, linewidth=.3),
        panel.background = element_rect(fill = "white"))

print(Map2)
```
```{r}
scale_factor = 10

## Save as PNG file
ggsave("Gifs/Map2_%03d.png",
       plot = last_plot(), 
       device = "png",
       scale = scale_factor,
       units="mm", 
       limitsize = FALSE,
       dpi=300)
```
```{r}
roads <- opq(bb) %>%
  add_osm_feature(key = "highway", value = c("motorway", "motorway_link")) %>%
  osmdata_sf()
```

```{r}
roadsConstruction <- opq(bb) %>%
  add_osm_feature(key = "highway", value = c("construction")) %>%
  osmdata_sf()

```

```{r}
examine <- roadsConstruction$osm_lines %>%
  dplyr::select(!c(starts_with(c("FIXME","abutters","access","adfc","area","bdouble","bicycle","bridge:", "building", "bus","class","change", "check_date:","cycleway",
                                 
#                    "communication","construction:","covered","cutting","d#estination:","disused", "ele","embankment","emergency","foot","hazard","hazmat","HGV","horse","lane#s","fixme", "int_ref:","junction", "lane","layer","level","max","min","operator:type", #"operator:w","height","incline","lit","description:","motor","mofa","name:","ref:","service","shoul#der","seamark","sidewalk","psv","overtaking","parking","loc","mapi","nat","oneway","placement","pla#nned","preserved", "priority","smoothness","start_","surface","survey","taxi","toll","traffic","tro#lley","tunnel:alt","tunnel:name:","turn:",
                    "surface","tactile","toilets","tracktype","url","vehicle","wheel","width"))))
 
 View(examine)

```


```{r message=FALSE, warning=FALSE}
toproads <- roads$osm_lines %>%
    dplyr::select(name, osm_id) %>%
    group_by(name) %>%
    tidyr::drop_na() %>%
    summarise(n = n()) %>%
    arrange(desc(n)) %>%
    head(30)
toproads
```
```{r message=FALSE, warning=FALSE}
Map2 +
geom_sf(data = toproads, size = .05, aes(colour = name))
```

```{r}
## Save as PNG file
ggsave("Gifs/Map_roads_%03d.png",
       plot = last_plot(), 
       device = "png",
       scale = scale_factor,
       units="mm", 
       limitsize = FALSE,
       dpi=300)
```
```{r}
topprimaryroads <- roads$osm_lines %>%
    dplyr::select(int_ref, ref, official_ref, name, osm_id) %>%
    dplyr::filter(!int_ref %in% c("E 04", "E 45", "E 39", "E 04;E 06;E 20", "E 55", "E 06;E 20", "E 20", "E 06", "E 26", "E 234")) %>%
    group_by(int_ref, ref, official_ref, name) %>%
    tidyr::drop_na() %>%
    summarise(n = n()) %>%
    arrange(desc(n)) %>%
    head(40)
topprimaryroads
```




```{r}
Map2 +
geom_sf(data = topprimaryroads, size = .05, aes(colour = ref))
```


```{r message=FALSE, warning=FALSE}
Map2 +
geom_sf(data = topprimaryroads, size = .05, aes(colour = int_ref))

```


```{r}
#Øresund Bridge
bridges <- opq(bb) %>%
  add_osm_feature(key = "man_made", value = c("bridge")) %>%
  osmdata_sf()

```

```{r}
routeA <- roads[["osm_lines"]] %>% 
  filter(int_ref %in% c("E 20;E 47;E 55", "E 47", "E 47;E 55", "E 22")) %>%
  filter(!ref == "A 20") 
```

```{r}
route <- roads[["osm_lines"]] %>% 
  filter(name %in% c("Sydmotorvejen", "Amagermotorvejen",
                     "Øresundsmotorvejen","Køge Bugt Motorvejen", "Tunnel Billwerder-Moorfleet")|
           int_ref == "E 22")
```



```{r}
route_Constructed <- roadsConstruction[["osm_lines"]] %>% 
  filter(name %in% c("Sydmotorvejen", "Amagermotorvejen",
                     "Øresundsmotorvejen","Køge Bugt Motorvejen", "Tunnel Billwerder-Moorfleet")|
           int_ref == "E 47")
```

```{r}
b <- route_Constructed %>%
    dplyr::select(osm_id, bridge, construction, int_ref,opening_date, ref, tunnel, `tunnel:name`, geometry)
View(b)
```

```{r}
route_Constructed_T <- roadsConstruction[["osm_lines"]] %>% 
  filter(name %in% c("Sydmotorvejen", "Amagermotorvejen",
                     "Øresundsmotorvejen","Køge Bugt Motorvejen", "Tunnel Billwerder-Moorfleet")|
           int_ref == "E 47") %>%
  filter(tunnel== "yes")
```



```{r}
Fehmarnbelt <- data.frame(pair = c("Fehmarnbelt","Fehmarnbelt"),
                    town = c("Puttgarden", "Rødbyhavn"),
                    lat = c(54.49676, 54.66587),
                    lon = c(11.211643, 11.344733))


Fehmarnbelt
```



```{r message=FALSE, warning=FALSE}
Map3 <- ggplot() +
    geom_sf(
    data = sea,
    fill = "aliceblue",
    color = "skyblue1", # outline
    linewidth=.2
    ) +
    geom_sf(data = roads$osm_lines, size = .05, colour = "lightgray") +
    geom_sf(data = primary_streets$osm_lines, size = .01, colour = "gray90") +
    geom_sf(data = Points, size = Points$PointSize,, shape = 21, fill = "darkred") +
    geom_sf_text(data = Labels, aes(label = Label, lineheight = .85), 
               size = Labels$FontSize,
               colour = Labels$FontColour,
               fontface = Labels$FontType,
               nudge_x = Labels$NudgeX,
               nudge_y = Labels$NudgeY) +
      geom_sf(data=route, colour = "navy", size = 1) +
      geom_sf(data=route_Constructed_T, colour = "darkred", size = .5, linetype="dotted") +
     # geom_line(data = Fehmarnbelt, 
     #       colour = "darkred", size = .5,
     #       linetype=2,
     #       aes(x = lon, y = lat, group = pair))+
  coord_sf(xlim = c(lon_min, lon_max), 
           ylim = c(lat_min, lat_max-2.5),
           expand = FALSE) +
    theme(axis.line = element_line(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
       # axis.ticks = element_blank(),  # hides the Lat, Long tick lines
       # axis.text.x = element_blank(), # hides the Longitude lines
       # axis.text.y = element_blank(), # hides the Latitude lines
        axis.title =element_blank(),  # hides the X and y axis labels
        panel.border = element_rect(colour = "black", fill=NA, linewidth=.3),
        panel.background = element_rect(fill = "white"))

print(Map3)
```




```{r}
Map3a <- ggplot() +
    geom_sf(
    data = sea,
    fill = "aliceblue",
    color = "skyblue1", # outline
    linewidth=.2
    ) +
    geom_sf(data = roads$osm_lines, size = .05, colour = "lightgray") +
    geom_sf(data = primary_streets$osm_lines, size = .01, colour = "gray90") +
    geom_sf(data = Points, size = Points$PointSize,, shape = 21, fill = "darkred") +
    geom_sf_text(data = Labels, aes(label = Label, lineheight = .85), 
               size = Labels$FontSize,
               colour = Labels$FontColour,
               fontface = Labels$FontType,
               nudge_x = Labels$NudgeX,
               nudge_y = Labels$NudgeY) +
      geom_sf(data=routeA, colour = "navy", size = 1) +
      geom_sf(data=filter(route_Constructed, !tunnel %in% "yes" | !construction %in% "motorway"), colour = "darkgrey", size = 2, linetype="solid") +
        geom_sf(data=filter(route_Constructed, tunnel== "yes" & construction=="motorway"), colour = "darkred", size = .5, linetype="dotted") +
     # geom_line(data = Fehmarnbelt, 
     #       colour = "darkred", size = .5,
     #       linetype=2,
     #       aes(x = lon, y = lat, group = pair))+
  coord_sf(xlim = c(lon_min, lon_max), 
           ylim = c(lat_min, lat_max-2.5),
           expand = FALSE) +
    theme(axis.line = element_line(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
       # axis.ticks = element_blank(),  # hides the Lat, Long tick lines
       # axis.text.x = element_blank(), # hides the Longitude lines
       # axis.text.y = element_blank(), # hides the Latitude lines
        axis.title =element_blank(),  # hides the X and y axis labels
        panel.border = element_rect(colour = "black", fill=NA, linewidth=.3),
        panel.background = element_rect(fill = "white"))

print(Map3a)

```


```{r}
scale_factor = 10

## Save as PNG file
ggsave("Gifs/Map3a_%03d.png",
       plot = last_plot(), 
       device = "png",
       scale = scale_factor,
       units="mm", 
       limitsize = FALSE,
       dpi=300)

```
```{r}
examine <- roads$osm_lines %>%
  dplyr::select(!c(FIXME, `abandoned:highway`,  starts_with(c("access","aeroway","bridge:","bus","change", "check_date:","communication","construction:","covered","cutting","cycleway","destination:","disused", "ele","embankment","emergency","foot","hazard","hazmat","HGV","horse","lanes","fixme", "int_ref:","junction", "lane","layer","level","max","min","operator:type", "operator:w","height","incline","lit","description:","motor","mofa","name:","ref:","service","shoulder","seamark","sidewalk","psv","overtaking","parking","loc","mapi","nat","oneway","placement","planned","preserved", "priority","smoothness","start_","surface","survey","taxi","toll","traffic","trolley","tunnel:alt","tunnel:name:","turn:","source:max","urban","vehicle","osm")),  agricultural, bdouble, bicycle))
View(examine)
glimpse(examine)
```


```{r message=FALSE, warning=FALSE}
toproadsByRef <- roads$osm_lines %>%
    dplyr::select(ref, int_ref, name, destination, osm_id) %>%
    group_by(ref,int_ref, name, destination) %>%
    tidyr::drop_na() %>%
    summarise(n = n()) %>%
    arrange(desc(n)) %>%
    head(40)
toproadsByRef
```
```{r message=FALSE, warning=FALSE}
Map2 +
geom_sf(data = toproadsByRef, size = .05, aes(colour = ref))

```

