---
title: "R Notebook"
output: html_notebook
---


```{r}
library(lwgeom)
library(osmdata)
library(sf)
library(dplyr)
library(ggplot2)
```

```{r}
### define example bbox
lon_min <- 6.5 # xmin
lon_max <- 13 # xmax
lat_min <- 53.5 # ymin
lat_max <- 58.5 # ymax

bb <- matrix(c(lon_min,lon_max,lat_min,lat_max), byrow = TRUE, nrow = 2, ncol = 2, dimnames = list(c('x','y'),c('min','max'))) 
bb

```

```{r}
### get "water" that is not sea as polygons
# water <- opq(bb) %>%
#   add_osm_feature(key = "natural", value = "water") %>%
#   osmdata_sf()
```


```{r}
### get sea & land as polygons
# 1. get coastline (as line)
coast <- opq(bb) %>%
  add_osm_feature(key = "natural", value = "coastline") %>%
  osmdata_sf()

```

```{r message=FALSE, warning=FALSE}
polygon_Map <- ggplot() +
  geom_sf(
    data = coast$osm_polygons,
    fill = "azure",
    color = "steelblue"
  ) +
  coord_sf(xlim = c(lon_min, lon_max), 
           ylim = c(lat_min, lat_max),
           expand = FALSE) +
    theme(axis.line = element_line(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=.3),
        panel.background = element_rect(fill = "white"))

print(polygon_Map)

scale_factor = 10

## Save as PNG file
ggsave("Gifs/Map1_Polygons_%03d.png",
       plot = last_plot(), 
       device = "png",
       scale = scale_factor,
       units="mm", 
       limitsize = FALSE,
       dpi=300)
```

```{r message=FALSE, warning=FALSE}
line_Map <- ggplot() +
  geom_sf(
    data = coast$osm_lines,
    fill = "azure",
    color = "steelblue",
    linewidth=.3
  ) + 
  coord_sf(xlim = c(lon_min, lon_max), 
           ylim = c(lat_min, lat_max),
           expand = FALSE) +
    theme(axis.line = element_line(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=.3),
        panel.background = element_rect(fill = "white"))

print(line_Map)

scale_factor = 10

## Save as PNG file
ggsave("Gifs/Map1_Line_%03d.png",
       plot = last_plot(), 
       device = "png",
       scale = scale_factor,
       units="mm", 
       limitsize = FALSE,
       dpi=300)

```


```{r message=FALSE, warning=FALSE}
line_and_polygon_Map <- ggplot() +
    geom_sf(
    data = coast$osm_polygons,
    fill = "red",
    color = "green"#,
  #  linewidth=.3
    ) +
  geom_sf(
    data = coast$osm_lines,
    fill = "azure",
    color = "steelblue",
    linewidth=.3
  ) + 
  coord_sf(xlim = c(lon_min, lon_max), 
           ylim = c(lat_min, lat_max),
           expand = FALSE) +
    theme(axis.line = element_line(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(colour = "black", fill=NA, linewidth=.3),
        panel.background = element_rect(fill = "white"))

print(line_and_polygon_Map)

scale_factor = 10

## Save as PNG file
ggsave("Gifs/Map1_Line_and_Polygon_%03d.png",
       plot = last_plot(), 
       device = "png",
       scale = scale_factor,
       units="mm", 
       limitsize = FALSE,
       dpi=300)

```

There are a small number of islets that are not included in the osm_lines. So merge all of the polygons (*osm_polygons*) into lines (*osm_lines*) with the function **osm_poly2line**.

```{r}
# Merge all of the osm_polygon objects in coast into osm_lines. The osm_polygons are retained but the are also repeated as osm_lines
coast_all <- osm_poly2line(coast)

dim(coast$osm_lines)
dim(coast_all$osm_lines)
```

However to create a map with sea and land in colour we need to convert osm_lines into polygons. 
This is done by creating a rectangle/polygon which is the size of the bounding box (bbox) and then spliting it via the coastline (**coast_all$osm_lines**). 
```{r}
# Get overall rectangle for bbox
bb_rect <- data.frame(
  lat = c(lat_min, lat_max),
  lon = c(lon_min, lon_max)
) %>% 
  st_as_sf(coords = c("lon", "lat"), crs = 4326) %>% 
  st_bbox() %>% 
  st_as_sfc()

```


```{r}
# Split overall rectangle for bbox via coastline
bb_rect_split <- bb_rect %>% 
  st_split(coast_all$osm_lines) %>% 
  st_collection_extract("POLYGON")
```

```{r}
# Extract the split parts
sea <- bb_rect_split[1]
```


```{r message=FALSE, warning=FALSE}
Map1 <- ggplot() +
    geom_sf(
    data = sea,
    fill = "aliceblue",
    color = "skyblue1", # outline
    linewidth=.3
    ) +
  coord_sf(xlim = c(lon_min, lon_max), 
           ylim = c(lat_min, lat_max),
           expand = FALSE) +
    theme(axis.line = element_line(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
       # axis.ticks = element_blank(),  # hides the Lat, Long tick lines
       # axis.text.x = element_blank(), # hides the Longitude lines
       # axis.text.y = element_blank(), # hides the Latitude lines
        panel.border = element_rect(colour = "black", fill=NA, linewidth=.3),
        panel.background = element_rect(fill = "white"))

print(Map1)

scale_factor = 10

## Save as PNG file
ggsave("Gifs/Map1_%03d.png",
       plot = last_plot(), 
       device = "png",
       scale = scale_factor,
       units="mm", 
       limitsize = FALSE,
       dpi=300)

```

```{r}
# Read in file containing towns, seas and countries
Map_coordinates <- readxl::read_excel("Data/Map_coordinates.xlsx", 
                           sheet = "Sheet1", range = "A1:D18",
                           col_types = c("text", "numeric", "numeric", "text")) %>%
    # Add a new column "Mark" which is set to Label if the Type is Sea or Country and otherwise set to Point
  dplyr::mutate(Mark = case_when(Type %in% c("Sea", "Country", "Area") ~ "Label",
                                 TRUE ~ "Point")) %>%
  # Add a new column "Fontsize" which is set to 4 if Type is Country, 
  # 3 if Sea and 2.5 if any other value
  dplyr::mutate(FontSize = case_when(
    Type == "Country" ~ 4,
    Type == "Sea" ~ 3,
    TRUE ~ 3))  %>%
  dplyr::mutate(FontType = case_when(
    Type == "Country" ~ "bold",
    Type == "Sea" ~ "bold.italic",
    Type == "Area" ~ "italic",
    TRUE ~ "plain"))  %>%
  dplyr::mutate(FontColour = case_when(
    Type == "Country" ~ "black",
    Label %in% c("Puttgarden","Rødbyhavn") ~ "black",
    Type == "Sea" ~ "steelblue",
    TRUE ~ "darkgray"))  %>%
    # Add a new column "NudgeX" which is set to 0 
  dplyr::mutate(NudgeX = case_when(
    Label %in% c("Puttgarden","Rødbyhavn") ~ -.33,
    Type == "Country" ~ 0,
    Type == "Sea" ~ 0,
    TRUE ~ 0)) %>%
    # Add a new column "NudgeX" which is set to 0 
  dplyr::mutate(NudgeY = case_when(
    Label %in% c("Puttgarden","Rødbyhavn") ~ 0,
    Type == "Country" ~ 0,
    Type == "Sea" ~ 0,
    TRUE ~ -.05))

  
head(Map_coordinates,3)
```

```{r}
Labels <- Map_coordinates %>%
  sf::st_as_sf(coords = c("Long", "Lat"), crs = 4326, agr = "constant")

Points <- dplyr::filter(Map_coordinates, !Mark == "Label") %>%
  sf::st_as_sf(coords = c("Long", "Lat"), crs = 4326, agr = "constant")

```



```{r message=FALSE, warning=FALSE}
Map2 <- ggplot() +
    geom_sf(
    data = sea,
    fill = "aliceblue",
    color = "skyblue1", # outline
    linewidth=.2
    ) +
    geom_sf(data = Points, size = 1, shape = 21, fill = "darkred") +
    geom_sf_text(data = Labels, aes(label = Label, lineheight = .85), 
               size = Labels$FontSize,
               colour = Labels$FontColour,
               fontface = Labels$FontType,
               nudge_x = Labels$NudgeX,
               nudge_y = Labels$NudgeY) +
  coord_sf(xlim = c(lon_min, lon_max), 
           ylim = c(lat_min, lat_max-2.5),
           expand = FALSE) +
    theme(axis.line = element_line(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
       # axis.ticks = element_blank(),  # hides the Lat, Long tick lines
       # axis.text.x = element_blank(), # hides the Longitude lines
       # axis.text.y = element_blank(), # hides the Latitude lines
        axis.title =element_blank(),  # hides the X and y axis labels
        panel.border = element_rect(colour = "black", fill=NA, linewidth=.3),
        panel.background = element_rect(fill = "white"))

print(Map2)
```
```{r}
scale_factor = 10

## Save as PNG file
ggsave("Gifs/Map2_%03d.png",
       plot = last_plot(), 
       device = "png",
       scale = scale_factor,
       units="mm", 
       limitsize = FALSE,
       dpi=300)
```
```{r}
roads <- opq(bb) %>%
  add_osm_feature(key = "highway", value = c("motorway", "motorway_link")) %>%
  osmdata_sf()
```

```{r}
roads2 <- opq(bb) %>%
  add_osm_feature(key = "highway", value = c("primary", "primary_link")) %>%
  osmdata_sf()
```



```{r}
library(tidyr)
```

```{r message=FALSE, warning=FALSE}
toproads <- roads$osm_lines %>%
    dplyr::select(name, osm_id) %>%
   # as.data.frame() %>%
    group_by(name) %>%
    tidyr::drop_na() %>%
    summarise(n = n()) %>%
    arrange(desc(n)) %>%
    head(20)
toproads
```
```{r message=FALSE, warning=FALSE}
Map2 +
geom_sf(data = toproads, size = .05, aes(colour = name))
```

```{r}
## Save as PNG file
ggsave("Gifs/Map_roads_%03d.png",
       plot = last_plot(), 
       device = "png",
       scale = scale_factor,
       units="mm", 
       limitsize = FALSE,
       dpi=300)
```


```{r}
NordjyskeMotorvej <- roads[["osm_lines"]] %>% 
  filter(name %in% c("Nordjyske Motorvej"))

SønderjyskeMotorvej <- roads[["osm_lines"]] %>% 
  filter(name %in% c("Sønderjyske Motorvej"))

FynskeMotorvej<- roads[["osm_lines"]] %>% 
  filter(name %in% c("Fynske Motorvej"))

Holbækmotorvejen <- roads[["osm_lines"]] %>% 
  filter(name %in% c("Holbækmotorvejen"))

ØstjyskeMotorvej <- roads[["osm_lines"]] %>% 
  filter(name %in% c("Østjyske Motorvej"))

Vestmotorvejen <- roads[["osm_lines"]] %>% 
  filter(name %in% c("Vestmotorvejen"))

Norgevägen <- roads[["osm_lines"]] %>% 
  filter(name %in% c("Norgevägen"))

Boråsleden <- roads[["osm_lines"]] %>% 
  filter(name %in% c("Boråsleden"))

MidtjyskeMotorvej <- roads[["osm_lines"]] %>% 
  filter(name %in% c("Midtjyske Motorvej"))

Helsingørmotorvejen <- roads[["osm_lines"]] %>% 
  filter(name %in% c("Helsingørmotorvejen"))

Hirtshalsmotorvejen <- roads[["osm_lines"]] %>% 
  filter(name %in% c("Hirtshalsmotorvejen"))

Lundbyleden <- roads[["osm_lines"]] %>% 
  filter(name %in% c("Lundbyleden"))

Sydmotorvejen <- roads[["osm_lines"]] %>% 
  filter(name %in% c("Sydmotorvejen"))

Motorring3 <- roads[["osm_lines"]] %>% 
  filter(name %in% c("Motorring 3"))
  
KøgeBugt <- roads[["osm_lines"]] %>% 
  filter(name %in% c("Køge Bugt Motorvejen"))

Amagermotorvejen <-roads[["osm_lines"]] %>% 
  filter(name %in% c("Amagermotorvejen"))
  
Øresundsmotorvejen <- roads[["osm_lines"]] %>% 
  filter(name %in% c("Øresundsmotorvejen"))

Trollhätteleden <- roads[["osm_lines"]] %>% 
  filter(name %in% c("Trollhätteleden"))



```
```{r}
#Øresund Bridge
bridges <- opq(bb) %>%
  add_osm_feature(key = "man_made", value = c("bridge")) %>%
  osmdata_sf()

```

```{r}
route <- roads[["osm_lines"]] %>% 
  filter(name %in% c("Sydmotorvejen", "Amagermotorvejen",
                     "Øresundsmotorvejen","Køge Bugt Motorvejen"))
```
```{r}
Fehmarnbelt <- data.frame(pair = c("Fehmarnbelt","Fehmarnbelt"),
                    town = c("Puttgarden", "Rødbyhavn"),
                    lat = c(54.49676, 54.66587),
                    lon = c(11.211643, 11.344733))


Fehmarnbelt
```



```{r message=FALSE, warning=FALSE}
Map3 <- ggplot() +
    geom_sf(
    data = sea,
    fill = "aliceblue",
    color = "skyblue1", # outline
    linewidth=.2
    ) +
    geom_sf(data = roads$osm_lines, size = .05, colour = "lightgray") +
    geom_sf(data = primary_streets$osm_lines, size = .01, colour = "gray90") +
    geom_sf(data = Points, size = 1, shape = 21, fill = "darkred") +
    geom_sf_text(data = Labels, aes(label = Label, lineheight = .85), 
               size = Labels$FontSize,
               colour = Labels$FontColour,
               fontface = Labels$FontType,
               nudge_x = Labels$NudgeX,
               nudge_y = Labels$NudgeY) +
    geom_line(data = Fehmarnbelt, 
            colour = "darkred", size = .5,
            linetype=2,
            aes(x = lon, y = lat, group = pair))+
      geom_sf(data=route, colour = "navy", size = 1) +
#  geom_sf(data=Lundbyleden, colour = "green", size = 1) +
#   geom_sf(data = MidtjyskeMotorvej, colour = "green", size = 1) +
#     geom_sf(data = Hirtshalsmotorvejen, colour = "purple", size = 1) +
#   geom_sf(data = Boråsleden, colour = "green", size = 1) +
#   geom_sf(data = Norgevägen, colour = "green", size = 1) +
#      geom_sf(data = Helsingørmotorvejen, colour = "navy", size = 1) +
#      geom_sf(data = Motorring3, colour = "purple", size = 1) +
#      geom_sf(data = Vestmotorvejen, colour = "green", size = 1) +
      #geom_sf(data = FynskeMotorvej, colour = "green", size = 1) +
  #Trollhätteleden
       # geom_sf(data = Trollhätteleden, colour = "green", size = 1) +
  #NordjyskeMotorvej
       # geom_sf(data = NordjyskeMotorvej, colour = "green", size = 1) +
  #Holbækmotorvejen
       # geom_sf(data = Holbækmotorvejen, colour = "green", size = 1) +
    #geom_sf(data = ØstjyskeMotorvej, colour = "green", size = 1) +
    #  geom_sf(data = SønderjyskeMotorvej, colour = "green", size = 1) +
  coord_sf(xlim = c(lon_min, lon_max), 
           ylim = c(lat_min, lat_max-2.5),
           expand = FALSE) +
    theme(axis.line = element_line(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
       # axis.ticks = element_blank(),  # hides the Lat, Long tick lines
       # axis.text.x = element_blank(), # hides the Longitude lines
       # axis.text.y = element_blank(), # hides the Latitude lines
        axis.title =element_blank(),  # hides the X and y axis labels
        panel.border = element_rect(colour = "black", fill=NA, linewidth=.3),
        panel.background = element_rect(fill = "white"))

print(Map3)
```






```{r message=FALSE, warning=FALSE}
Map4 <- Map3 +
  geom_line(data = Fehmarnbelt, 
            colour = "darkred", size = .5,
            linetype=2,
            aes(x = lon, y = lat, group = pair))

print(Map4)
```



```{r}
scale_factor = 10

## Save as PNG file
ggsave("Gifs/Map3_%03d.png",
       plot = last_plot(), 
       device = "png",
       scale = scale_factor,
       units="mm", 
       limitsize = FALSE,
       dpi=300)

```

